Semi-Permeable Vesicles: Mathematical Formulation, Numerics, and Physics

### 2020-12-11 vesicle_code/docs/permeableVes/main.tex
edits to results

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Quiescent Flow} 
Our analysis indicates that a semi-permeable vesicle in a quiescent flow
relaxes to a circle with constant tension $\Lambda = 2\pi^2/L^2$.
Furthermore, equation~\eqref{eqn:areaROC} shows that the area change is
proportional to $\beta$, and the times for the vesicle to deviate from
its initial reduced area and to reach its steady state should both be
proportional to $\beta^{-1}$. This is verified for a vesicle with
initial reduced area $\nu = 0.65$ and with different semi-permeability
rates. In Table~\ref{tbl:ellipseRelaxTimes} we report the required time,
$t_\mathrm{steady}$, for the vesicle's reduced area to be within 1.3\%
of its steady state reduced area, and the required time,
$t_\mathrm{inflate}$, for the vesicle's reduced area to exceed 1.3\% of
its initial reduced area. Both times are proportional to $\beta^{-1}$,
and for this reason, our numerical results are reported in terms of the
scaled time $\beta t$.
\begin{table}[htp]
  \centering
  \caption{\label{tbl:ellipseRelaxTimes} The time for a semi-permeable
  vesicle to its steady state ($t_\mathrm{steady}$) and to begin
  inflating ($t_\mathrm{inflate}$).}
  \begin{tabular}{C{0.8cm}C{2cm}C{0.8cm}C{2cm}C{0.8cm}}
    $\beta$ & $t_\mathrm{steady}$  & ratio 
            & $t_\mathrm{inflate}$ & ratio \\
    \midrule
    $10^{0}$  & $6.45 \times 10^{0}$ & ---  
              & $2.17 \times 10^{-2}$ & --- \\ 
    $10^{-1}$ & $2.00 \times 10^{1}$ & 3.10 & 
                $1.25 \times 10^{-1}$ & 5.76 \\
    $10^{-2}$ & $9.61 \times 10^{1}$ & 4.81 & 
                $1.59 \times 10^{0}$ & 12.72 \\
    $10^{-3}$ & $8.36 \times 10^{2}$ & 8.71 & 
                $2.19 \times 10^{1}$ & 13.77 \\
    $10^{-4}$ & $8.15 \times 10^{3}$ & 9.74 & 
                $2.14 \times 10^{2}$ & 9.77 \\
    $10^{-5}$ & $8.03 \times 10^{4}$ & 9.85 & 
                $2.20 \times 10^{3}$ & 10.14 \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/relaxationComposite.tikz}
  \fi
  \caption{\label{fig:relaxationComposite} (a): Snapshots of four
  semi-permeable vesicles with $\beta = 10^{-3}$ submerged in a
  quiescent flow. The fluid flux along the red regions is into the
  vesicle, and the fluid flux along the blue regions is out of the
  vesicle. In all cases, the vesicle reaches a steady-state circular
  shape. (b): The reduced area of each of the simulations. The
  particular snapshots in (a) occur at the marks along the curve.}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Shear Flow}
\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/shearComposite.tikz}
  \fi
  \caption{\label{fig:shearComposite} (a) The reduced area of a
  semi-permeable vesicle with $\beta = 10^{-3}$ initialized with four
  different reduced areas in a shear flow with two different flow rates.
  (b) The steady state shape of a semi-permeable vesicle with varying
  flow rates and semi-permeability rates. The circled vesicles
  correspond to the simulations in part (a).}
\end{figure*}

We consider a semi-permeable vesicle in the shear flow
$\uu_{\infty}(\xx) = \dot{\gamma} (y,0)$ with different permeability
rates $\beta$ and shear rates $\dot{\gamma}$. A vesicle's reduced area
and viscosity contrast determines whether it tank treads or
tumbles~\cite{fin-lam-sei-gom2008, kra-win-sei-lip1996}. Similar to the
dynamics of an impermeable vesicle, a semi-permeable vesicle tilts to an
inclination angle and undergoes tank treading dynamics. Since we
consider semi-permeable vesicles, there is no viscosity contrast and
tumbling is not possible.

In Figure~\ref{fig:shearComposite}(a), using four initial reduced areas
and two flow rates, we plot the reduced area of a semi-permeable vesicle
with $\beta = 10^{-3}$. As observed in a quiescent flow, the final
vesicle shape is independent of the initial reduced area. However, it
does depend on the flow rate. We also observe small changes in the
reduced area at early times with the vesicle dynamics at early times
that are similar to those of an impermeable vesicle. However, once the
vesicle is tank-treading, the normal component of the force-balance
velocity is small and the permeability velocity dominates and leads to a
changing reduced area.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Poiseuille Flow}
\begin{figure*}[htp]
%  \ifTikz
  \input{figures/parabolicComposite.tikz}
%  \fi
  \caption{\label{fig:parabolicComposite} (a) The steady state shape of a
  semi-permeable vesicle submerged in a Poiseuille flow with varying
  initial reduced areas and flow rates. (b) The steady-state inclination
  angle for four different flow velocities. The steady-state shapes are
  superimposed. (c) The reduced area of the slipper formed with the
  flow rate of 800 $\mu m/s$. (d) The vesicle shape shown with the
  background imposed flow. The channel width is 12.5 times larger than
  the vesicle radius. The corresponding reduced area for each shape is
  indicated by the marks in plot (c).}
  \todo[inline]{shift right plot to the right and increase necessary
  font sizes}
\end{figure*}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Quiescent Flow} 
\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/relaxationComposite.tikz}
  \fi
  \caption{\label{fig:relaxationComposite} (a): Snapshots of four
  semi-permeable vesicles with $\beta = 10^{-3}$ submerged in a
  quiescent flow. The fluid flux along the red regions is into the
  vesicle, and the fluid flux along the blue regions is out of the
  vesicle. In all cases, the vesicle reaches a steady state circular
  shape. (b): The reduced area of each of the simulations. The
  particular snapshots in (a) occur at the marks along the curve.}
\end{figure*}
Our analysis shows that a semi-permeable vesicle in a quiescent flow
relaxes to a circle at a rate that is proportional to $\beta$, and the
steady state tension is $\Lambda = 2\pi^2/L^2$. Therefore, the times for
the vesicle to deviate from its initial reduced area and to reach its
steady state are proportional to $\beta^{-1}$. This is verified for a
vesicle with initial reduced area $\nu = 0.65$ and with different
semi-permeability rates. In Table~\ref{tbl:ellipseRelaxTimes} we report
the required time, $t_\mathrm{steady}$, for the vesicle's reduced area
to be within 1.3\% of its steady state reduced area, and the required
time, $t_\mathrm{inflate}$, for the vesicle's reduced area to exceed
1.3\% of its initial reduced area. Both times are proportional to
$\beta^{-1}$, and for this reason, our numerical results are reported in
terms of the scaled time $\beta t$.
\begin{table}[htp]
  \centering
  \caption{\label{tbl:ellipseRelaxTimes} The time for a semi-permeable
  vesicle to its steady state ($t_\mathrm{steady}$) and to begin
  inflating ($t_\mathrm{inflate}$).}
  \begin{tabular}{C{0.8cm}C{2cm}C{0.8cm}C{2cm}C{0.8cm}}
    $\beta$ & $t_\mathrm{steady}$  & ratio 
            & $t_\mathrm{inflate}$ & ratio \\
    \midrule
    $10^{0}$  & $6.45 \times 10^{0}$ & ---  
              & $2.17 \times 10^{-2}$ & --- \\ 
    $10^{-1}$ & $2.00 \times 10^{1}$ & 3.10 & 
                $1.25 \times 10^{-1}$ & 5.76 \\
    $10^{-2}$ & $9.61 \times 10^{1}$ & 4.81 & 
                $1.59 \times 10^{0}$ & 12.72 \\
    $10^{-3}$ & $8.36 \times 10^{2}$ & 8.71 & 
                $2.19 \times 10^{1}$ & 13.77 \\
    $10^{-4}$ & $8.15 \times 10^{3}$ & 9.74 & 
                $2.14 \times 10^{2}$ & 9.77 \\
    $10^{-5}$ & $8.03 \times 10^{4}$ & 9.85 & 
                $2.20 \times 10^{3}$ & 10.14 \\
    \bottomrule
  \end{tabular}
\end{table}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Shear Flow}
\begin{figure}[htp]
  \centering
  \ifTikz
  \input{figures/shearComposite.tikz}
  \fi
  \caption{\label{fig:shearComposite} (a) The reduced area of a
  semi-permeable vesicle with $\beta = 10^{-3}$ initialized with four
  different reduced areas in a shear flow with two different flow rates.
  (b) The steady state shape of a semi-permeable vesicle with varying
  flow rates and semi-permeability rates. The circled vesicles
  correspond to the simulations in part (a).}
\end{figure}

We consider a semi-permeable vesicle in the shear flow
$\uu_{\infty}(\xx) = \dot{\gamma} (y,0)$ with different permeability
rates $\beta$ and shear rates $\dot{\gamma}$. A vesicle's reduced area
and viscosity contrast determine whether it tank treads or
tumbles~\cite{fin-lam-sei-gom2008, kra-win-sei-lip1996}. Similar to the
dynamics of an impermeable vesicle, a semi-permeable vesicle tilts to an
inclination angle and undergoes tank treading dynamics. Since we
consider semi-permeable vesicles, there is no viscosity contrast and
we do not observe tumbling.

In Figure~\ref{fig:shearComposite}(a), using four initial reduced areas
and two flow rates, we plot the reduced area of a semi-permeable vesicle
with $\beta = 10^{-3}$. As observed in a quiescent flow, the final
vesicle shape is independent of the initial reduced area. However, it
does depend on the flow rate. We also observe small changes in the
reduced area at early times when the semi-permeable vesicle behaves like
an impermeable vesicle. However, once the vesicle is tank-treading, the
normal component of the force-balance velocity is small and the
permeability velocity dominates and leads to a changing reduced area.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Poiseuille Flow}
\begin{figure*}[htp]
  \ifTikz
  \centering
  \input{figures/parabolicComposite.tikz}
  \fi
  \caption{\label{fig:parabolicComposite} (a) The steady state shape of
  a semi-permeable vesicle submerged in a Poiseuille flow with varying
  initial reduced areas and flow rates. (b) The steady state inclination
  angle for four different flow velocities. The steady state shapes are
  superimposed. (c) The reduced area of the slipper formed with the flow
  rate of 800 $\mu m/s$. (d) The vesicle shape shown with the background
  imposed flow. The channel width is 12.5 times larger than the vesicle
  radius. The corresponding reduced area of each vesicle is indicated by
  the marks in plot (c).}
\end{figure*}
